---
output: github_document
---

# {{{package}}}

Describe your {{{package}}} here.

## How to install:

We’ve made it easier than ever to install and start analysing global governance data in R. Simply install the core package, [manydata](https://github.com/globalgov/manydata), as outlined below, to discover, install, and update various packages from the many packages universe.

```{r get, warning=FALSE, message=FALSE}
manydata::get_packages() # this prints a list of the publicly available data packages currently available
#manydata::get_packages("{{{package}}}") # this downloads and installs the named package
```

## Data included

Once you have installed '{manydata}', you can see the different databases and datasets included in the package using the following function.

```{r datasets, warning=FALSE, message=FALSE}
manydata::data_contrast("{{{package}}}")
```

Working with ensembles of related data has many advantages for robust
analysis. Just take a look at our vignettes
[here](https://globalgov.github.io/manydata/articles/user.html).

While some of our packages can and do include novel data,
much of what they offer involves standing on the shoulders of giants.
Our packages endeavour to be as transparent as possible about where data comes from,
how it has been coded and/or relabeled, and who has done the work.
As such, we make it easy to cite both the particular datasets you use by listing their
original references with the first function below, as well as the package
providers for their work assembling the data using the second function.

``` {r cites}
# Citing the original data
manydata::data_source("{{{package}}}")
# Citing the package
citation("{{{package}}}")
```

## Contributing

[manydata](https://github.com/globalgov/manydata) and [manypkgs](https://github.com/globalgov/manypkgs) also makes it easy to contribute in lots of different ways.

If you have already developed a dataset salient to this package,
please reach out by flagging it as an
[issue](https://github.com/globalgov/{{{package}}}/issues) for us, or by
forking, further developing the package yourself, and opening a [pull
request](https://github.com/globalgov/{{{package}}}/pulls) so that your data
can easily be integrated into the package.

If you have collected or developed other data that may not thematically fit
in this package,
but could be useful within the many packages universe,
[manypkgs](https://github.com/globalgov/manypkgs) includes a number of
functions that make it easy to create a new package and populate with global governance data.

If you have any further ideas about how this package, or the various other 'many packages',
might better facilitate your empirical analysis,
we’d be very happy to hear from you.
